Certified Knowledge Compilation \\ with Application to Formally Verified Model Counting

### 2023-11-14 doc/JAIR/jair-paper.tex
Some improvements

\definecolor{redorange}{rgb}{0.878431, 0.235294, 0.192157}
\definecolor{lightblue}{rgb}{0.552941, 0.72549, 0.792157}
\definecolor{clearyellow}{rgb}{0.964706, 0.745098, 0}
\definecolor{clearorange}{rgb}{0.917647, 0.462745, 0}
\definecolor{mildgray}{rgb}{0.54902, 0.509804, 0.47451}
\definecolor{softblue}{rgb}{0.643137, 0.858824, 0.909804}
\definecolor{bluegray}{rgb}{0.141176, 0.313725, 0.603922}
\definecolor{lightgreen}{rgb}{0.709804, 0.741176, 0}
\definecolor{darkgreen}{rgb}{0.152941, 0.576471, 0.172549}
\definecolor{redpurple}{rgb}{0.835294, 0, 0.196078}
\definecolor{midblue}{rgb}{0, 0.592157, 0.662745}
\definecolor{clearpurple}{rgb}{0.67451, 0.0784314, 0.352941}
\definecolor{browngreen}{rgb}{0.333333, 0.313725, 0.145098}
\definecolor{darkestpurple}{rgb}{0.396078, 0.113725, 0.196078}
\definecolor{greypurple}{rgb}{0.294118, 0.219608, 0.298039}
\definecolor{darkturquoise}{rgb}{0, 0.239216, 0.298039}
\definecolor{darkbrown}{rgb}{0.305882, 0.211765, 0.160784}
\definecolor{midgreen}{rgb}{0.560784, 0.6, 0.243137}
\definecolor{darkred}{rgb}{0.576471, 0.152941, 0.172549}
\definecolor{darkpurple}{rgb}{0.313725, 0.027451, 0.470588}
\definecolor{darkestblue}{rgb}{0, 0.156863, 0.333333}
\definecolor{lightpurple}{rgb}{0.776471, 0.690196, 0.737255}
\definecolor{softgreen}{rgb}{0.733333, 0.772549, 0.572549}
\definecolor{offwhite}{rgb}{0.839216, 0.823529, 0.768627}
\definecolor{medgreen}{rgb}{0.15, 0.6, 0.15}

@@

@@

A CPOG file provides both a declaration of a POG, as well as a checkable
proof that a Boolean formula, given in conjunctive normal
form, is logically equivalent to the POG\@.
The proof format draws its inspiration from the LRAT~\cite{lrat} and
QRAT~\cite{heule:JAR2014} formats for unquantified and quantified Boolean formulas, respectively.
Key properties include:
\begin{itemize}
  \item
  The file contains declarations of $\pand$ and $\por$ operations to describe the POG.
  Declaring a node $\nodeu$ implicitly adds an extension variable $u$ and a set of defining clauses $\theta_{u}$
  encoding the product or sum operation.
  This is the only means for adding extension variables to the proof.
\item Boolean negation is supported implicitly by allowing the
  arguments of the $\por$ and $\pand$ operations to be literals and not just
  variables.
\item
  The file contains explicit clause addition steps.
  A clause can only be added if it is logically implied by the existing clauses.
  A sequence of clause identifiers must be listed as a \emph{hint} providing a RUP verification of the implication.
\item
  The file contains explicit clause deletion steps.
  A clause can only be deleted if it is logically implied by the remaining clauses.
  A sequence of clause identifiers must be listed as a \emph{hint} providing a RUP verification of the implication.
\item The checker must track the dependency set for every input and
  extension variable.  For each $\pand$ operation, the checker must ensure that the dependency sets for its arguments are disjoint.
  The associated extension variable has a dependency set equal to the union of those of its arguments.
\item Declaring a $\por$ operation requires a sequence of clauses
  providing a RUP proof that the arguments are mutually exclusive.
  Only binary $\por$ operations are allowed to avoid requiring multiple proofs of disjointness
%  Generalizing to $k$ arguments would require
%  $k\,(k-1)/2$ proofs of disjointness.
\end{itemize}

@@

1 & \texttt{-1 3 -4} & \texttt{0} \\
2 & \texttt{-1 -3 4} & \texttt{0} \\
3 & \texttt{3 -4} & \texttt{0}\\
4 & \texttt{1 -3 4} & \texttt{0} \\
5 & \texttt{-1 -2} & \texttt{0} \\

@@

6 & \texttt{p 5 -3 -4} & \texttt{0} & $p_5 = \obar{x}_3 \pand \obar{x}_4$ \\
9 & \texttt{p 6 3 4} & \texttt{0} & $p_6 = x_3 \pand x_4$ \\
12 & \texttt{s 7 5 6 7 10} & \texttt{0} & $s_7 = p_5 \por p_6$ \\
15 & \texttt{p 8 -1 7} & \texttt{0} & $p_8 = \obar{x}_1 \pand s_7$ \\
18 & \texttt{p 9 1 -2 7} & \texttt{0} & $p_9 = x_1 \pand \obar{x}_2 \pand s_7$ \\
22 & \texttt{s 10 8 9 16 19} & \texttt{0} & $s_{10} = p_8 \por p_9$ \\

@@

\texttt{6} & \texttt{5 3 4} & \texttt{0} & Define $p_5$ \\
\texttt{7} & \texttt{-5 -3} & \texttt{0} & \\
\texttt{8} & \texttt{-5 -4} & \texttt{0} & \\

@@

\texttt{9} & \texttt{6 -3 -4} & \texttt{0} & Define $p_6$ \\
\texttt{10} & \texttt{-6 3} & \texttt{0} & \\
\texttt{11} & \texttt{-6 4} & \texttt{0} & \\

@@

\texttt{12} & \texttt{-7 5 6} & \texttt{0} & Define $s_7$ \\
\texttt{13} & \texttt{7 -5} & \texttt{0} & \\
\texttt{14} & \texttt{7 -6} & \texttt{0} & \\

@@

\texttt{15} & \texttt{8 1 -7} & \texttt{0} & Define $p_8$ \\
\texttt{16} & \texttt{-8 -1} & \texttt{0} & \\
\texttt{17} & \texttt{-8 7} & \texttt{0} & \\

@@

\texttt{18} & \texttt{9 -1 2 -7} & \texttt{0} & Define $p_9$ \\
\texttt{19} & \texttt{-9 1} & \texttt{0} & \\
\texttt{20} & \texttt{-9 -2} & \texttt{0} & \\
\texttt{21} & \texttt{-9 7} & \texttt{0} & \\

@@

\texttt{22} & \texttt{-10 8 9} & \texttt{0} & Define $s_{10}$ \\
\texttt{23} & \texttt{10 -8} & \texttt{0} & \\
\texttt{24} & \texttt{10 -9} & \texttt{0} & \\

@@

\texttt{25} & \texttt{a 5 1 3} & \texttt{0} & \texttt{3 6} & \texttt{0} & $\obar{x}_1 \land \obar{x}_3 \imply p_5$ \\
\texttt{26} & \texttt{a 6 1 -3} & \texttt{0} & \texttt{4 9} & \texttt{0} & $\obar{x}_1 \land x_3 \imply p_6$ \\
\texttt{27} & \texttt{a 3 7 1} & \texttt{0} & \texttt{13 25} & \texttt{0} & $\obar{x}_3 \land \obar{x}_1 \imply s_7$  \\
\texttt{28} & \texttt{a 7 1} & \texttt{0} & \texttt{27 14 26} & \texttt{0} & $\obar{x}_1 \imply s_7$  \\
\texttt{29} & \texttt{a 8 1} & \texttt{0} & \texttt{28 15} & \texttt{0} & $\obar{x}_1 \imply p_8$  \\
\texttt{30} & \texttt{a 5 -1 3} & \texttt{0} & \texttt{1 6} & \texttt{0} & $x_1 \land \obar{x}_3 \imply p_5$ \\
\texttt{31} & \texttt{a 6 -1 -3} & \texttt{0} & \texttt{2 9} & \texttt{0} & $x_1 \land x_3 \imply p_6$ \\
\texttt{32} & \texttt{a 3 7 -1} & \texttt{0} & \texttt{13 30} & \texttt{0} & $\obar{x}_3 \land x_1 \imply s_7$  \\
\texttt{33} & \texttt{a 7 -1} & \texttt{0} & \texttt{32 14 31} & \texttt{0} & $x_1 \imply s_7$  \\
\texttt{34} & \texttt{a 9 -1} & \texttt{0} & \texttt{5 33 18} & \texttt{0} & $x_1 \imply p_9$  \\
\texttt{35} & \texttt{a 1 10} & \texttt{0} & \texttt{23 29} & \texttt{0} & $\obar{x}_1 \imply s_{10}$  \\
\texttt{36} & \texttt{a 10} & \texttt{0} & \texttt{35 24 34} & \texttt{0} & $s_{10}$ \\

@@

 \texttt{d 1} & \texttt{36 8 10 12 16 21 22} & \texttt{0} & Delete clause 1 \\
 \texttt{d 2} & \texttt{36 7 11 12 16 21 22} & \texttt{0} & Delete clause 2 \\
 \texttt{d 3} & \texttt{36 8 10 12 17 19 22} & \texttt{0} & Delete clause 3 \\
 \texttt{d 4} & \texttt{36 7 11 12 17 19 22} & \texttt{0} & Delete clause 4 \\
 \texttt{d 5} & \texttt{36 16 20 22} & \texttt{0} &  Delete clause 5 \\

@@

\caption{Example formula (A), its POG representation (B), and its CPOG proof (C), (E), and (F)}

@@

We step through portions of the file to provide a better understanding of the CPOG proof framework.
Figure
\ref{fig:eg4:all}(D) shows the defining clauses that are implicitly
defined by the POG operation declarations.  These do not appear in the
CPOG file.  Referring back to the declarations of the sum nodes in
Figure \ref{fig:eg4:all}(C), we can see that the declaration of node
$\nodes_7$ had clause IDs 7 and 10 as the hint.  We can see in Figure
\ref{fig:eg4:all}(A) that these two clauses form a RUP proof for the clause
$\obar{p}_5 \lor \obar{p}_6$, showing that the two children of $\nodes_7$
have disjoint models.  Similarly, node $\nodes_{10}$ is declared as having
clause IDs 16 and 19 as the hint.  These form a RUP proof for the clause
$\obar{p}_8 \lor \obar{p}_9$, showing that the two children of
$\nodes_{10}$ have disjoint models.

@@

Figure \ref{fig:eg4:all}(E) provides the sequence of assertions
leading to unit clause 36, consisting of the literal $s_{10}$.  This clause indicates that $\nodes_{10}$ is implied by the input clauses, i.e.,
any total assignment $\assign$
satisfying the input clauses must have its extension to $\eassign$ yield $\eassign(s_{10}) = 1$.
Working backward, we can see that
clause 29 indicates that variable $p_8$ will be implied by the input
clauses when $\assign(x_1) = 0$, while clause 34 indicates that node $p_9$ will
be implied by the input clauses when $\assign(x_1) = 1$.  These serve as the
hint for clause 36.

@@

Deleting the asserted clauses is so simple that we do not show it.  It
involves simply deleting the clauses from clause number 35 down to
clause number 25, with each deletion using the same hint as were used
to add that clause.  In the end, therefore, only the defining clauses
for the POG nodes and the unit clause asserting $s_{10}$ remain,
completing a proof that the POG is logically equivalent to the input
formula.

@@

The performance of the structural proof generator for forward implication, both in its execution time and
the size of the proof generated, can be improved by two optimizations
described here.  A key feature is that they do not require any changes
to the proof framework---they build on the power of extended
resolution to enable new logical structures.  They
involve declaring new product nodes to encode products of literals.
These nodes are not part of the POG
representation of the formula; they serve only to enable the forward
implication proof.

@@

\begin{figure}
(A) Additional nodes\\[1.0em]
\begin{tabular}{llll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{CPOG line} & Explanation \\
\midrule
25 & \texttt{p 11 -3 4} & \texttt{0} & $v_{11} = \obar{x}_3 \pand {x}_4$ \\
28 & \texttt{p 12  3 -4} & \texttt{0} & $v_{12} = {x}_3 \pand \obar{x}_4$ \\
\bottomrule
\end{tabular}
\\[1.0em]
(B) Implicit Clauses\\[1.2em]
\begin{tabular}{llll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{Clauses} & Explanation \\
\midrule
\texttt{25} & \texttt{11 3 -4} & \texttt{0} & Argument clause $\{x_3 ,\, \obar{x}_4\}$, activated by $\obar{v}_{11}$ \\
\texttt{26} & \texttt{-11 -3} & \texttt{0} & \\
\texttt{27} & \texttt{-11 4} & \texttt{0} & \\
\midrule
\texttt{28} & \texttt{12 -3 4} & \texttt{0} & Argument clause $\{\obar{x}_3,\,  {x}_4\}$, activated by $\obar{v}_{12}$ \\
\texttt{29} & \texttt{-12 3} & \texttt{0} & \\
\texttt{30} & \texttt{-12 -4} & \texttt{0} & \\
\bottomrule
\end{tabular}
\\[1.0em]
(C) CPOG Assertions\\[1.0em]
\begin{tabular}{llllll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{Clause} & \multicolumn{2}{l}{Hint} & Explanation \\
\midrule
\multicolumn{6}{l}{Lemma Proof} \\
\texttt{31} & \texttt{a 5 11 12 3} & \texttt{0} & \texttt{25 6} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land \obar{x}_3 \imply p_{5}$ \\
\texttt{32} & \texttt{a 6 11 12 -3} & \texttt{0} & \texttt{28 9} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land {x}_3 \imply p_{6}$ \\
\texttt{33} & \texttt{a 3 7 11 12} & \texttt{0} & \texttt{13 31} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land \obar{x}_3 \imply s_{7}$ \\
\texttt{34} & \texttt{a 7 11 12} & \texttt{0} & \texttt{33 14 32} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \imply s_{7}$ \\
\midrule
\multicolumn{6}{l}{Lemma Application \#1} \\
\texttt{35} & \texttt{a -11 1} & \texttt{0} & \texttt{26 27 3} & \texttt{0} & $\obar{x}_1 \imply \obar{v}_{11}$ \\
\texttt{36} & \texttt{a -12 1} & \texttt{0} & \texttt{29 30 4} & \texttt{0} & $\obar{x}_1 \imply \obar{v}_{12}$ \\
\texttt{37} & \texttt{a 7 1} & \texttt{0} & \texttt{35 36 34} & \texttt{0} & $\obar{x}_1 \imply s_7$ \\
\midrule
\texttt{38} & \texttt{a 8 1} & \texttt{0} & \texttt{37 15} & \texttt{0} & $\obar{x}_1 \imply p_8$ \\
\midrule
\multicolumn{6}{l}{Lemma Application \#2} \\
\texttt{39} & \texttt{a -11 -1} & \texttt{0} & \texttt{26 27 1} & \texttt{0} & ${x}_1 \imply \obar{v}_{11}$ \\
\texttt{40} & \texttt{a -12 -1} & \texttt{0} & \texttt{29 30 2} & \texttt{0} & ${x}_1 \imply \obar{v}_{12}$ \\
\texttt{41} & \texttt{a 7 -1} & \texttt{0} & \texttt{39 40 34} & \texttt{0} & ${x}_1 \imply s_7$ \\
\midrule
\texttt{42} & \texttt{a 9 -1} & \texttt{0} & \texttt{5 41 18} & \texttt{0} & ${x}_1 \imply p_9$ \\
\texttt{43} & \texttt{a 1 10} & \texttt{0} & \texttt{23 38} & \texttt{0} & $\obar{x}_1 \imply s_{10}$  \\
\texttt{44} & \texttt{a 10} & \texttt{0} & \texttt{43 24 42} & \texttt{0} & $s_{10}$ \\
\bottomrule
\end{tabular}
\caption{Example of lemma definition, proof, and application}
\label{fig:eg4:lemmas}
\end{figure}

Figure~\ref{fig:eg4:lemmas} shows an alternate forward implication
proof for the example of Figure~\ref{fig:eg4:all} using a lemma to
represent the shared node $\nodes_7$.  We can see that the POG with
this node as root encodes the Boolean formula $x_3 \leftrightarrow x_4$, having a CNF representation consisting of the clauses
$\{x_3 ,\, \obar{x}_4\}$ and $\{\obar{x}_3 ,\, {x}_4\}$.  The product node
declarations shown in Figure~\ref{fig:eg4:lemmas}(A) create synthetic
clauses 25 and 28 to encode these arguments with activating literals
$\obar{v}_{11}$ and $\obar{v}_{12}$, respectively.  Clauses 31--34
then provide a proof of the lemma, stating that any assignment
$\assign$ that activates these clauses will  assign $1$ to $s_7$.
Clauses 35 and 36 state that an assignment with $\assign(x_1) = 0$
will cause the first synthetic clause to activate due to input clause
3, and it will cause the second synthetic clause to activate due to
input clause 4.  From this, clause 37 can use the lemma to state that
assigning $0$ to $x_1$ will cause $s_7$ to evaluate to $1$.  Similarly,
clauses 39 and 40 serve to activate the synthetic clauses when
$\assign(x_1) = 1$, due to input clauses 1 and 2, and clause 41 then
uses the lemma to state that assigning $1$ to $x_1$ will cause $s_7$ to
evaluate to $1$.

>>>>

\definecolor{redorange}{rgb}{0.878431, 0.235294, 0.192157}
\definecolor{lightblue}{rgb}{0.552941, 0.72549, 0.792157}
\definecolor{clearyellow}{rgb}{0.964706, 0.745098, 0}
\definecolor{clearorange}{rgb}{0.917647, 0.462745, 0}
\definecolor{mildgray}{rgb}{0.54902, 0.509804, 0.47451}
\definecolor{softblue}{rgb}{0.643137, 0.858824, 0.909804}
\definecolor{bluegray}{rgb}{0.141176, 0.313725, 0.603922}
\definecolor{lightgreen}{rgb}{0.709804, 0.741176, 0}
\definecolor{darkgreen}{rgb}{0.152941, 0.576471, 0.172549}
\definecolor{redpurple}{rgb}{0.835294, 0, 0.196078}
\definecolor{midblue}{rgb}{0, 0.592157, 0.662745}
\definecolor{clearpurple}{rgb}{0.67451, 0.0784314, 0.352941}
\definecolor{browngreen}{rgb}{0.333333, 0.313725, 0.145098}
\definecolor{darkestpurple}{rgb}{0.396078, 0.113725, 0.196078}
\definecolor{greypurple}{rgb}{0.294118, 0.219608, 0.298039}
\definecolor{darkturquoise}{rgb}{0, 0.239216, 0.298039}
\definecolor{darkbrown}{rgb}{0.305882, 0.211765, 0.160784}
\definecolor{midgreen}{rgb}{0.560784, 0.6, 0.243137}
\definecolor{darkred}{rgb}{0.576471, 0.152941, 0.172549}
\definecolor{darkpurple}{rgb}{0.313725, 0.027451, 0.470588}
\definecolor{darkestblue}{rgb}{0, 0.156863, 0.333333}
\definecolor{lightpurple}{rgb}{0.776471, 0.690196, 0.737255}
\definecolor{softgreen}{rgb}{0.733333, 0.772549, 0.572549}
\definecolor{offwhite}{rgb}{0.839216, 0.823529, 0.768627}
\definecolor{medgreen}{rgb}{0.15, 0.6, 0.15}
\definecolor{midred}{rgb}{0.80,0.3,0.3}

@@

\newcommand{\rtext}[1]{\textcolor{midred}{\texttt{#1}}}
\newcommand{\btext}[1]{\textcolor{midblue}{\texttt{#1}}}

@@

A CPOG file provides both a declaration of a POG, as well as a checkable
proof that a Boolean formula, given in conjunctive normal
form, is logically equivalent to the POG\@.
The proof format draws its inspiration from the LRAT~\cite{lrat} and
QRAT~\cite{heule:JAR2014} formats for unquantified and quantified Boolean formulas, respectively.
Key properties include:
\begin{itemize}
  \item
  The file contains declarations of $\pand$ and $\por$ operations to describe the POG.
  Declaring a node $\nodeu$ implicitly adds an extension variable $u$ and a set of defining clauses $\theta_{u}$
  encoding the product or sum operation.
  This is the only means for adding extension variables to the proof.
\item Boolean negation is supported implicitly by allowing the
  arguments of the $\por$ and $\pand$ operations to be literals and not just
  variables.
\item
  The file contains explicit clause addition steps.
  A clause can only be added if it is logically implied by the existing clauses.
  A sequence of clause identifiers must be listed as a \emph{hint} providing a RUP verification of the implication.
\item
  The file contains explicit clause deletion steps.
  A clause can only be deleted if it is logically implied by the remaining clauses.
  A sequence of clause identifiers must be listed as a \emph{hint} providing a RUP verification of the implication.
\item The checker must track the dependency set for every input and
  extension variable.  For each $\pand$ operation, the checker must ensure that the dependency sets for its arguments are disjoint.
  The associated extension variable has a dependency set equal to the union of those of its arguments.
\item Declaring a $\por$ operation requires a sequence of clauses
  providing a RUP proof that the arguments are mutually exclusive.
  Only binary $\por$ operations are allowed to avoid requiring multiple proofs of disjointness.
%  Generalizing to $k$ arguments would require
%  $k\,(k-1)/2$ proofs of disjointness.
\end{itemize}

@@

\rtext{1} & \texttt{-1 3 -4} & \texttt{0} \\
\rtext{2} & \texttt{-1 -3 4} & \texttt{0} \\
\rtext{3} & \texttt{3 -4} & \texttt{0}\\
\rtext{4} & \texttt{1 -3 4} & \texttt{0} \\
\rtext{5} & \texttt{-1 -2} & \texttt{0} \\

@@

\rtext{6} & \texttt{p 5 -3 -4} & \texttt{0} & $p_5 = \obar{x}_3 \pand \obar{x}_4$ \\
\rtext{9} & \texttt{p 6 3 4} & \texttt{0} & $p_6 = x_3 \pand x_4$ \\
\rtext{12} & \texttt{s 7 5 6} \; \rtext{7 10} & \texttt{0} & $s_7 = p_5 \por p_6$ \\
\rtext{15} & \texttt{p 8 -1 7} & \texttt{0} & $p_8 = \obar{x}_1 \pand s_7$ \\
\rtext{18} & \texttt{p 9 1 -2 7} & \texttt{0} & $p_9 = x_1 \pand \obar{x}_2 \pand s_7$ \\
\rtext{22} & \texttt{s 10 8 9} \; \rtext{16 19} & \texttt{0} & $s_{10} = p_8 \por p_9$ \\

@@

\rtext{6} & \texttt{5 3 4} & \texttt{0} & Define $p_5$ \\
\rtext{7} & \texttt{-5 -3} & \texttt{0} & \\
\rtext{8} & \texttt{-5 -4} & \texttt{0} & \\

@@

\rtext{9} & \texttt{6 -3 -4} & \texttt{0} & Define $p_6$ \\
\rtext{10} & \texttt{-6 3} & \texttt{0} & \\
\rtext{11} & \texttt{-6 4} & \texttt{0} & \\

@@

\rtext{12} & \texttt{-7 5 6} & \texttt{0} & Define $s_7$ \\
\rtext{13} & \texttt{7 -5} & \texttt{0} & \\
\rtext{14} & \texttt{7 -6} & \texttt{0} & \\

@@

\rtext{15} & \texttt{8 1 -7} & \texttt{0} & Define $p_8$ \\
\rtext{16} & \texttt{-8 -1} & \texttt{0} & \\
\rtext{17} & \texttt{-8 7} & \texttt{0} & \\

@@

\rtext{18} & \texttt{9 -1 2 -7} & \texttt{0} & Define $p_9$ \\
\rtext{19} & \texttt{-9 1} & \texttt{0} & \\
\rtext{20} & \texttt{-9 -2} & \texttt{0} & \\
\rtext{21} & \texttt{-9 7} & \texttt{0} & \\

@@

\rtext{22} & \texttt{-10 8 9} & \texttt{0} & Define $s_{10}$ \\
\rtext{23} & \texttt{10 -8} & \texttt{0} & \\
\rtext{24} & \texttt{10 -9} & \texttt{0} & \\

@@

\rtext{25} & \texttt{a 5 1 3} & \texttt{0} & \rtext{3 6} & \texttt{0} & $\obar{x}_1 \land \obar{x}_3 \imply p_5$ \\
\rtext{26} & \texttt{a 6 1 -3} & \texttt{0} & \rtext{4 9} & \texttt{0} & $\obar{x}_1 \land x_3 \imply p_6$ \\
\rtext{27} & \texttt{a 3 7 1} & \texttt{0} & \rtext{13 25} & \texttt{0} & $\obar{x}_3 \land \obar{x}_1 \imply s_7$  \\
\rtext{28} & \texttt{a 7 1} & \texttt{0} & \rtext{27 14 26} & \texttt{0} & $\obar{x}_1 \imply s_7$  \\
\rtext{29} & \texttt{a 8 1} & \texttt{0} & \rtext{28 15} & \texttt{0} & $\obar{x}_1 \imply p_8$  \\
\rtext{30} & \texttt{a 5 -1 3} & \texttt{0} & \rtext{1 6} & \texttt{0} & $x_1 \land \obar{x}_3 \imply p_5$ \\
\rtext{31} & \texttt{a 6 -1 -3} & \texttt{0} & \rtext{2 9} & \texttt{0} & $x_1 \land x_3 \imply p_6$ \\
\rtext{32} & \texttt{a 3 7 -1} & \texttt{0} & \rtext{13 30} & \texttt{0} & $\obar{x}_3 \land x_1 \imply s_7$  \\
\rtext{33} & \texttt{a 7 -1} & \texttt{0} & \rtext{32 14 31} & \texttt{0} & $x_1 \imply s_7$  \\
\rtext{34} & \texttt{a 9 -1} & \texttt{0} & \rtext{5 33 18} & \texttt{0} & $x_1 \imply p_9$  \\
\rtext{35} & \texttt{a 1 10} & \texttt{0} & \rtext{23 29} & \texttt{0} & $\obar{x}_1 \imply s_{10}$  \\
\rtext{36} & \texttt{a 10} & \texttt{0} & \rtext{35 24 34} & \texttt{0} & $s_{10}$ \\

@@

 \texttt{d 1} & \rtext{36 8 10 12 16 21 22} & \texttt{0} & Delete clause 1 \\
 \texttt{d 2} & \rtext{36 7 11 12 16 21 22} & \texttt{0} & Delete clause 2 \\
 \texttt{d 3} & \rtext{36 8 10 12 17 19 22} & \texttt{0} & Delete clause 3 \\
 \texttt{d 4} & \rtext{36 7 11 12 17 19 22} & \texttt{0} & Delete clause 4 \\
 \texttt{d 5} & \rtext{36 16 20 22} & \texttt{0} &  Delete clause 5 \\

@@

\caption{Example formula (A), its POG representation (B), and its CPOG proof (C), (E), and (F).  The defining clauses (D) are implicitly
defined by the POG declaration (C).}

@@

We step through portions of the file to provide a better understanding of the CPOG proof framework.
Figure
\ref{fig:eg4:all}(D) shows the defining clauses that are implicitly
defined by the POG operation declarations.  These do not appear in the
CPOG file.  Referring back to the declarations of the sum nodes in
Figure \ref{fig:eg4:all}(C), we can see that the declaration of node
$\nodes_7$ has clause IDs 7 and 10 as the hint.  We can see in Figure
\ref{fig:eg4:all}(A) that these two clauses form a RUP proof for the clause
$\obar{p}_5 \lor \obar{p}_6$, showing that the two children of $\nodes_7$
have disjoint models.  Similarly, node $\nodes_{10}$ is declared as having
clause IDs 16 and 19 as the hint.  These form a RUP proof for the clause
$\obar{p}_8 \lor \obar{p}_9$, showing that the two children of
$\nodes_{10}$ have disjoint models.

@@

Figure \ref{fig:eg4:all}(E) provides the sequence of assertions
leading to unit clause 36, consisting of the literal $s_{10}$.  This clause indicates that $\nodes_{10}$ is implied by the input clauses, i.e.,
any total assignment $\assign$
satisfying the input clauses must have its extension to $\eassign$ yield $\eassign(s_{10}) = 1$.
Working backward, we can see that
clause 35 indicates that variable $s_{10}$ will be implied by the input
clauses when $\assign(x_1) = 0$. Clause 34 indicates that node $p_9$ will
be implied by the input clauses when $\assign(x_1) = 1$, while defining clause 24 shows that node $s_{10}$ will be implied
by the input clauses when $\eassign{p_9} = 1$.  These three clauses
serve as the
hint for clause 36.

@@

Deleting the asserted clauses is so simple that we do not show it.  It
involves simply deleting the clauses from clause number 35 down to
clause number 25, with each deletion using the same hint as was used
to add that clause.  In the end, therefore, only the defining clauses
for the POG nodes and the unit clause asserting $s_{10}$ remain,
completing a proof that the POG is logically equivalent to the input
formula.

@@

The performance of the structural proof generator for forward implication, both in its execution time and
the size of the proof generated, can be improved by two optimizations
described here.  A key feature is that they do not require any changes
to the proof framework---they build on the power of extended
resolution to enable the construction of new logical structures.  They
involve declaring new product nodes to encode products of literals.
These nodes are not part of the POG
representation of the formula; they serve only to enable the forward
implication proof.

@@

\begin{figure}
(A) Additional nodes\\[1.0em]
\begin{tabular}{llll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{CPOG line} & Explanation \\
\midrule
\rtext{25} & \texttt{p 11 -3 4} & \texttt{0} & $v_{11} = \obar{x}_3 \pand {x}_4$ \\
\rtext{28} & \texttt{p 12  3 -4} & \texttt{0} & $v_{12} = {x}_3 \pand \obar{x}_4$ \\
\bottomrule
\end{tabular}
\\[1.0em]
(B) Implicit Clauses\\[1.2em]
\begin{tabular}{llll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{Clauses} & Explanation \\
\midrule
\rtext{25} & \texttt{11 3 -4} & \texttt{0} & Argument clause $\{x_3 ,\, \obar{x}_4\}$, activated by $\obar{v}_{11}$ \\
\rtext{26} & \texttt{-11 -3} & \texttt{0} & \\
\rtext{27} & \texttt{-11 4} & \texttt{0} & \\
\midrule
\rtext{28} & \texttt{12 -3 4} & \texttt{0} & Argument clause $\{\obar{x}_3,\,  {x}_4\}$, activated by $\obar{v}_{12}$ \\
\rtext{29} & \texttt{-12 3} & \texttt{0} & \\
\rtext{30} & \texttt{-12 -4} & \texttt{0} & \\
\bottomrule
\end{tabular}
\\[1.0em]
(C) CPOG Assertions\\[1.0em]
\begin{tabular}{llllll}
\toprule
\makebox[5mm]{ID} & \multicolumn{2}{l}{Clause} & \multicolumn{2}{l}{Hint} & Explanation \\
\midrule
\multicolumn{6}{l}{Lemma Proof} \\
\rtext{31} & \texttt{a 5 11 12 3} & \texttt{0} & \rtext{25 6} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land \obar{x}_3 \imply p_{5}$ \\
\rtext{32} & \texttt{a 6 11 12 -3} & \texttt{0} & \rtext{28 9} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land {x}_3 \imply p_{6}$ \\
\rtext{33} & \texttt{a 3 7 11 12} & \texttt{0} & \rtext{13 31} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \land \obar{x}_3 \imply s_{7}$ \\
\rtext{34} & \texttt{a 7 11 12} & \texttt{0} & \rtext{33 14 32} & \texttt{0} & $(\obar{v}_{11} \land \obar{v}_{12}) \imply s_{7}$ \\
\midrule
\multicolumn{6}{l}{Lemma Application \#1} \\
\rtext{35} & \texttt{a -11 1} & \texttt{0} & \rtext{26 27 3} & \texttt{0} & $\obar{x}_1 \imply \obar{v}_{11}$ \\
\rtext{36} & \texttt{a -12 1} & \texttt{0} & \rtext{29 30 4} & \texttt{0} & $\obar{x}_1 \imply \obar{v}_{12}$ \\
\rtext{37} & \texttt{a 7 1} & \texttt{0} & \rtext{35 36 34} & \texttt{0} & $\obar{x}_1 \imply s_7$ \\
\midrule
\rtext{38} & \texttt{a 8 1} & \texttt{0} & \rtext{37 15} & \texttt{0} & $\obar{x}_1 \imply p_8$ \\
\midrule
\multicolumn{6}{l}{Lemma Application \#2} \\
\rtext{39} & \texttt{a -11 -1} & \texttt{0} & \rtext{26 27 1} & \texttt{0} & ${x}_1 \imply \obar{v}_{11}$ \\
\rtext{40} & \texttt{a -12 -1} & \texttt{0} & \rtext{29 30 2} & \texttt{0} & ${x}_1 \imply \obar{v}_{12}$ \\
\rtext{41} & \texttt{a 7 -1} & \texttt{0} & \rtext{39 40 34} & \texttt{0} & ${x}_1 \imply s_7$ \\
\midrule
\rtext{42} & \texttt{a 9 -1} & \texttt{0} & \rtext{5 41 18} & \texttt{0} & ${x}_1 \imply p_9$ \\
\rtext{43} & \texttt{a 1 10} & \texttt{0} & \rtext{23 38} & \texttt{0} & $\obar{x}_1 \imply s_{10}$  \\
\rtext{44} & \texttt{a 10} & \texttt{0} & \rtext{43 24 42} & \texttt{0} & $s_{10}$ \\
\bottomrule
\end{tabular}
\caption{Example of lemma definition, proof, and application}
\label{fig:eg4:lemmas}
\end{figure}

Figure~\ref{fig:eg4:lemmas} shows an alternate forward implication
proof for the example of Figure~\ref{fig:eg4:all} using a lemma to
represent the shared node $\nodes_7$.  We can see that the POG with
this node as root encodes the Boolean formula $x_3 \leftrightarrow x_4$, having a CNF representation consisting of the clauses
$\{x_3 ,\, \obar{x}_4\}$ and $\{\obar{x}_3 ,\, {x}_4\}$.  The product node
declarations shown in Figure~\ref{fig:eg4:lemmas}(A) create synthetic
clauses 25 and 28 to encode these arguments with activating literals
$\obar{v}_{11}$ and $\obar{v}_{12}$, respectively.  Clauses 31--34
then provide a proof of the lemma, stating that any assignment
$\assign$ that activates these clauses will, when extended,  assign $1$ to $s_7$.
Clauses 35 and 36 state that an assignment with $\assign(x_1) = 0$
will, when extended, cause the first synthetic clause to activate due to input clause
3, and it will cause the second synthetic clause to activate due to
input clause 4.  From this, clause 37 can use the lemma to state that
assigning $0$ to $x_1$ will cause $s_7$ to evaluate to $1$.  Similarly,
clauses 39 and 40 serve to activate the synthetic clauses when
$\assign(x_1) = 1$, due to input clauses 1 and 2, and clause 41 then
uses the lemma to state that assigning $1$ to $x_1$ will cause $s_7$ to
evaluate to $1$.
